---
title: "Options for grouping sites"
author: "Katie Renwick"
date: ""
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Possibly ways to group sites
1. Sagegrouse management units
2. Elevation within management unit
3. Elevation classes
4. Original climate groupings from Andy's PCA
5. Ecotypes (US Level 3 or NA Level 2)
6. Kuchler PNV types
7. Any other ideas???


```{r, echo=FALSE, include=F}
# intro code to read in and organize data
rm(list=ls())
library(ggplot2); theme_set(theme_bw(base_size=20)) # sized for ppt
library(gridExtra)
library(splines)
library(colorspace)
library(rgdal)
library(dplyr) # must load ggplot first!
library(tidyr)
library(knitr)
library(xtable)

# paths to data and folder for figures
dpath <- "/Users/poulterlab1/Box Sync/sageseer/ModelComparison/"
fpath <- "/Users/poulterlab1/Documents/sageseer/Figures/"
opath <- "/Users/poulterlab1/Documents/sageseer/"

# Pull in merged data and manipulate
merged <- read.csv(paste(dpath, "merged_data-co2.csv", sep=""))
m3 <- merged %>%
  mutate(change=predicted-baseline) %>%
  mutate(direction=ifelse(change>0,"Positive","Negative")) %>%
  na.omit() %>%
  dplyr::select(site,model:predicted,change,direction)
unit <- read.csv(paste(dpath, "focal_sites_by_zone.csv", sep="")) 

du <- merge(m3,unit, by="site") 

# Function to add sample size to boxplots:
give.n <- function(x){
  return(c(y = median(x)*3, label = length(x))) 
  # experiment with the multiplier (3) to find the perfect position
}
```

## 1. Sagegrouse management units
- Note sites not well distributed- ex. small sample in unit 6
- Some sites aren't in a management zone at all
- bird-centric, may not split sagebrush response in informative manner

```{r, echo = F,  results='asis',message=F}
# Code to map management units and show where the points fall
# Include table with sample size in each Mgmt Unit
d2 <- filter(du, var=="temp", mag==4)
t <- table(d2$model,d2$Mgmt_zone, useNA = "ifany")
TS <- t[1,]
Others <- t[3,]
nums <- rbind(TS,Others)
print(xtable(nums), scalebox='0.75',comment=F)
```

## Mgmt units: modeled response to 4C temperature increase

```{r, echo = F, fig.width = 15, fig.height = 10,  results='asis',message=F}
################################################################################
# Make bar plots showing change within each Mgmt. unit
################################################################################
d <- dplyr::filter(du, mag==4 & var=="temp")
ggplot(data=d, aes(x=Mgmt_zone,y=change, fill=model)) +
    geom_boxplot(notch=T) +
    ylab("Change in Response") +
  theme(panel.background=element_blank(),plot.background=element_blank(),
    legend.text.align = 0) +
  geom_hline(yintercept=0, linetype="dashed") #+
 # scale_y_continuous(limits=c(-50,50)) +
  #stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               #position = position_dodge(width = 0.75)) # label with sample size

```

## Elevation within mgmt unit
- problem: not enough sites to split into elev categories, also not a strong response to elev

```{r, echo = F, fig.width = 14, fig.height = 10,  results='asis',message=F}
# look at scatter- does elevation matter?
d <- dplyr::filter(du, mag==4 & var=="temp")
ggplot(data=d, aes(x=elev,y=change, color=model)) +
  geom_point() +
  geom_smooth(method="lm") +
  geom_hline(yintercept=0,linetype="dashed") +
  facet_wrap(~Name)
# ultimately, mgmt zone doesn't seem like a good classifier. Based on mgmt/
# sage grouse, so not surprising that doesn't work well for a plant

```

## Elevation bands alone

```{r, echo = F, fig.width = 14, fig.height = 10,  results='asis',message=F}
d <- dplyr::filter(du, mag==4 & var=="temp") %>%
  mutate(elev_cat=cut(elev, breaks=c(0,1000,1500,2000,2500,3000,4000), 
                      include.lowest=TRUE, labels=c("<1k","1k-1.5k","1.5k-2k","2k-2.5k","2.5k-3k",">3k")))
ggplot(data=d, aes(x=elev_cat,y=change, color=model)) +
  geom_boxplot(notch=F) +
  ylab("Change in Response") +
  xlab("Elevation Bin (m)") +
  theme(panel.background=element_blank(),plot.background=element_blank(),
        legend.text.align = 0) +
  geom_hline(yintercept=0, linetype="dashed") +
  stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75))

```

## Bands could be better. Look at full data set:
- all except DGVM suggest a cut-off between 1000-2000m
- elevation is really a proxy though- I'd rather stick with the MAT gradient if managers are okay with it.

```{r, echo = F, fig.width = 14, fig.height = 10,  results='asis',message=F}
# at this point, I'd rather just stick to the MAT plots
d <- dplyr::filter(du, mag==4 & var=="temp") 
ggplot(data=d, aes(x=elev,y=change, color=model)) +
  geom_point() +
  geom_smooth(method="lm") +
  ylab("Change in Response") +
  theme(panel.background=element_blank(),plot.background=element_blank(),
        legend.text.align = 0) +
  geom_hline(yintercept=0, linetype="dashed") +
  facet_wrap(~model, scales="free")

```

## Original climate groupings from Andy's PCA

```{r, echo = F, fig.width = 14, fig.height = 10,  results='asis',message=F}
d <- dplyr::filter(du, mag==4 & var=="temp") 
ggplot(data=d, aes(x=climate,y=change, color=model)) +
  geom_boxplot(notch=F) +
  ylab("Change in Response") +
  theme(panel.background=element_blank(),plot.background=element_blank(),
        legend.text.align = 0) +
  geom_hline(yintercept=0, linetype="dashed") +
  stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## What if I split these by ring? Not showing core.

```{r, echo = F, fig.width = 14, fig.height = 10,  results='asis',message=F}
d <- dplyr::filter(du, mag==4 & var=="temp" & climate!="core")
ggplot(data=d, aes(x=rings,y=change, color=model)) +
  geom_boxplot(notch=F) +
  ylab("Change in Response") +
  theme(panel.background=element_blank(),plot.background=element_blank(),
        legend.text.align = 0) +
  geom_hline(yintercept=0, linetype="dashed") +
  stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~climate)
```

## US Level 3 Ecoregions: sites per region

```{r, echo = F, fig.width = 14, fig.height = 10,  results='asis',message=F}
# Include table with sample size in each Mgmt Unit
d2 <- dplyr::filter(du, var=="temp", mag==4)
t <- table(d2$model,d2$US_L3NAME, useNA = "ifany")
TS <- t[1,]
Others <- t[3,]
nums <- cbind.data.frame(TS,Others)
nums2 <- nums[order(nums$Others),]
print(xtable(nums2), scalebox='0.75',comment=F)
```

## US Level 3 Ecoregions: Issues
- I eliminated the ecoregions with < 15 observations per model
- this eliminates a lot, and there are still too many categories. Can I lump them based on response? Proximity? Ecological characteristics?

## Sites grouped by ecoregion

```{r, echo = F, fig.width = 14, fig.height = 10,  results='asis',message=F}
d <- dplyr::filter(du, mag==4 & var=="temp") 
# which ecoregions do I have enough obs for?
d2 <-d %>% group_by(model,US_L3NAME) %>% mutate(n=n()) %>% filter(n>=15) %>% ungroup()
ggplot(data=d2, aes(x=US_L3NAME,y=change, color=model)) +
  geom_boxplot(notch=F) +
  ylab("Change in Response") +
  theme(panel.background=element_blank(),plot.background=element_blank(),
        legend.text.align = 0) +
  geom_hline(yintercept=0, linetype="dashed") +
  stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

## Sites grouped by L2 ecoregion (categories more general)

```{r, echo = F, fig.width = 14, fig.height = 10,  results='asis',message=F}
d <- dplyr::filter(du, mag==4 & var=="temp")
ggplot(data=d, aes(x=NA_L2NAME,y=change, color=model)) +
  geom_boxplot(notch=F) +
  ylab("Change in Response") +
  theme(panel.background=element_blank(),plot.background=element_blank(),
        legend.text.align = 0) +
  geom_hline(yintercept=0, linetype="dashed") +
  stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

## Sites grouped by Kuchler PNV types: sample size

```{r, echo = F, fig.width = 14, fig.height = 10,  results='asis',message=F}
d2 <- dplyr::filter(du, var=="temp", mag==4)
t <- table(d2$model,d2$PNV_GROUPS, useNA = "ifany")
TS <- t[1,]
Others <- t[3,]
nums <- cbind.data.frame(TS,Others)
nums2 <- nums[order(nums$Others),]
print(xtable(nums2), scalebox='0.75',comment=F)

```

## Sites grouped by Kuchler PNV types

```{r, echo = F, fig.width = 14, fig.height = 10,  results='asis',message=F}
d <- dplyr::filter(du, mag==4 & var=="temp")
ggplot(data=d, aes(x=PNV_GROUPS,y=change, color=model)) +
  geom_boxplot(notch=F) +
  ylab("Change in Response") +
  theme(panel.background=element_blank(),plot.background=element_blank(),
        legend.text.align = 0) +
  geom_hline(yintercept=0, linetype="dashed") +
  stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

## Sites grouped by Kuchler PNV types: lumped into categories (shapefile from Peter?)

```{r, echo = F, fig.width = 14, fig.height = 10,  results='asis',message=F}
d <- dplyr::filter(du, mag==4 & var=="temp")
ggplot(data=d, aes(x=VEGTYP_LAB,y=change, color=model)) +
  geom_boxplot(notch=F) +
  ylab("Change in Response") +
  theme(panel.background=element_blank(),plot.background=element_blank(),
        legend.text.align = 0) +
  geom_hline(yintercept=0, linetype="dashed") +
  stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Sites grouped by resistance and resilience classes

```{r, echo = F, fig.width = 14, fig.height = 10,  results='asis',message=F}
d <- dplyr::filter(du, mag==4 & var=="temp")
# cut sites with zero or NA
d2 <- filter(d, RR_class_name!=0) %>% na.omit()
ggplot(data=d2, aes(x=RR_class_name,y=change, color=model)) +
  geom_boxplot(notch=F) +
  ylab("Change in Response") +
  theme(panel.background=element_blank(),plot.background=element_blank(),
        legend.text.align = 0) +
  geom_hline(yintercept=0, linetype="dashed") +
  stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Sites grouped by temp/moisutre regime

```{r, echo = F, fig.width = 14, fig.height = 10,  results='asis',message=F}
d <- dplyr::filter(du, mag==4 & var=="temp")
# cut sites with zero or NA
d2 <- filter(d, RR_class_name!=0) %>% na.omit()
ggplot(data=d, aes(x=common_name,y=change, color=model)) +
  geom_boxplot(notch=F) +
  ylab("Change in Response") +
  theme(panel.background=element_blank(),plot.background=element_blank(),
        legend.text.align = 0) +
  geom_hline(yintercept=0, linetype="dashed") +
  stat_summary(fun.data = give.n, geom = "text", fun.y = median,
               position = position_dodge(width = 0.75)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```